Measuring precise radial velocities and cross-correlation function line-profile variations using a Skew Normal density\thanks{Based on observations collected at the La Silla Parana Observatory, ESO (Chile), with the HARPS spectrograph at the 3.6-m telescope.}

### 2018-07-17 manuscript.tex
updated Eq.~\ref{ to Eq.~\eqref{

The Skew Normal (SN) distribution is a class of probability distributions which includes the Normal distribution as a special case \citep{Azzalini1985}. The SN distribution has, in addition to a location and a scale parameter analogous to the Normal distribution's mean and standard deviation, a third parameter which describes the skewness (i.e. the asymmetry) of the distribution. Considering a random variable $Y\in \mathbb R$ (where $\mathbb R$ is the real line) which follows a SN distribution with location parameter $\xi \in \mathbb R$, scale parameter $\omega \in \mathbb R^{+}$ (i.e., the positive real line), and skewness parameter $\alpha \in \mathbb R$, its density at some value $y\in Y$ can be written as 
\begin{equation} \label{def:snd_gen}
SN(y;\xi, \omega, \alpha) = \frac{2}{\omega} \phi\left(\frac{y-\xi}{\omega}\right) \Phi\left(\frac{\alpha(y-\xi)}{\omega}\right),
\end{equation}
where $\phi$ and $\Phi$ are, respectively, the density function and the distribution function of a standard Normal distribution\footnote{A standard Normal distribution is a Normal distribution with a mean of 0 and a standard deviation of 1.}.
The skewness parameter $\alpha$ quantifies the asymmetry of the SN. 
Examples of SN densities under different skewness parameter values and the same location and scale parameters ($\xi = 0$ and $\omega = 1$) are displayed in Fig.~\ref{fig:SN.plot}.  A usual Normal distribution is the special case of the SN distribution when the skewness parameter $\alpha$ is equal to zero\footnote{This can be seen from Eq.~\ref{def:snd_gen}. If $\alpha = 0$ then $\Phi\left(\frac{\alpha(y-\xi)}{\omega}\right) = \Phi(0) = 0.5$ and therefore $SN(y;\xi, \omega, 0) = \frac{1}{\omega} \phi\left(\frac{y-\xi}{\omega}\right)$ which is the density of a Normal distribution. Note that $\Phi(0) = 0.5$ because $\Phi(0)$ is the the probability that a standard Normal random variable is less than or equal than 0.}.
%
\begin{figure}[t]
\begin{center}
\includegraphics[height = 2.3in]{Skew_Normal_densities_jjck.pdf} 
   \caption{Density function of a random variable $Y$ following the SN distribution $SN(\xi, \omega, \alpha)$ with location parameter $\xi = 0$, scale parameter $\omega = 1$ and different values of the skewness parameter $\alpha$ indicated by different colors and line types. Note that the solid black line has an $\alpha = 0$ making it a Normal distribution.}
   \label{fig:SN.plot}
\end{center}
\end{figure}
%
For reasons related to the interpretation of the parameters in Eq.~\ref{def:snd_gen} and computational issues with estimating $\alpha$ near 0, a different parametrization is used in this work, which is referred to as the \emph{centered parametrization} (CP).  This CP is much closer to the parametrization of a Normal distribution, as it uses a mean parameter $\mu$, a variance parameter $\sigma^2$ and a skewness parameter $\gamma$. In order to define the CP, we need to express the CP parameters $(\mu, \sigma^2, \gamma)$ as a function of $(\xi, \omega^2, \alpha)$. This can be done using the following relations:
%
\begin{equation} \label{eq:snd_cp}
\mu = \xi + \omega \beta, \quad \sigma^{2} = \omega^{2}(1-\beta^2), \quad \gamma = \frac{1}{2}(4-\pi) \beta^{3}\left(1-\beta^2\right)^{-3/2},
\end{equation}
%
where $\beta = \sqrt{\frac{2}{\pi}}\left(\frac{\alpha}{\sqrt{1+\alpha^2}}\right)$ \citep[e.g.][]{Arellano-2010}.

By using Eq.~\ref{eq:snd_cp}, the new set of parameters $(\mu, \sigma^2, \gamma)$ provides a clearer interpretation of the behavior of the SN distribution. For the $\alpha$ values used in Fig.~\ref{fig:SN.plot}, the corresponding values of ($\mu$, $\sigma^2$, $\gamma$) are displayed in Table~\ref{tab:cp_values}.  In particular, $\mu$ and $\sigma^2$ are the actual mean and variance of the distribution, rather than simply a location and scale parameter, and $\gamma$ provides a measure of the skewness of the SN. 
Along with the mean of the SN, we consider the median of the distribution as a measure of its barycenter.  See Table~\ref{tab:cp_values} for the medians of the SN densities displayed in Fig.~\ref{fig:SN.plot}.

@@

\end{center}
\end{table}
%
Further details about the parametrization from Eq.~\ref{def:snd_gen}, called the \emph{Direct Parametrization} (DP), the CP, and general statistical properties of the SN can be found in \cite{Azzalini2014}.

@@

The proposed function defined in Eq.~\ref{eq:RV:correction} is the result of statistical and astronomical considerations. 
In particular, we considered issues with correlations between the model covariates, which can lead to issues because the matrix needed to calculate the estimates could be singular; this problem is known in statistics as \emph{multicollinearity} and more discussion can be found in \citet{belsley1991}.
In the analysis of real data presented in this work, we never observed a correlation coefficient exceeding $0.66$ between the asymmetry and width parameters and therefore the problem of multicollinearity appeared to be mitigated. 
We also investigated the statistical significant of the interactions between $A$ and the width of the CCF, and $A$ and the asymmetry of the CCF; those two interactions were not statistically significant.

%-----------------------------------------------------------------------------------------------------------------------------------------------
\section{Simulation Study} \label{sec:soap}
In order to evaluate the performance of the proposed SN approach for modelling the CCF and the benefit of using the proposed correction for stellar activity (see Eq.~\ref{eq:RV:correction}), we begin by considering a simulation study using spectra generated from the Spot Oscillation And Planet 2.0 code \citep[SOAP 2.0,][]{Dumusque-2014b}.

@@

The RV variations displayed in Fig.~\ref{fig:faculae} are caused only by stellar activity, in this case a facula, so we applied the activity correction proposed in Eq.~\ref{eq:RV:correction} to evaluate the performance of the model in this setting.
The results of this correction are displayed in Fig.~\ref{fig:faculae.correction} and the statistical tests on the coefficients involved in Eq.~\ref{eq:RV:correction} are summarized in Table \ref{table:faculae.test}. 
The proposed correction for stellar activity is able to account for the majority of the activity signal created by a facula, with a $R^2$ larger than $0.99$. 
In addition, the rms of the different estimates of the RV reduces from about 3 \ms before correcting for stellar activity to values below 0.15 \ms after correcting for stellar activity.
We see a slightly smaller rms after correction when using the SN parameters compared to the Normal parameters.
When comparing the correction proposed in Eq.~\ref{eq:RV:correction} with what is generally used (i.e. a linear combination of only the asymmetry and width parameters), we see that the proposed correction is able to reduce the rms of the RV residuals by a factor of 2. Looking at the significance of the coefficients in Table~\ref{table:faculae.test}, we observe that all the SN or Normal parameters are relevant for the correction.

\begin{figure*}[htbp]
\begin{center}
\includegraphics[height = 6in]{FACULAE_NEW_CORRECTION_[3]CorrectionActivity_RadialVelocity_vs_time.pdf} 
   \caption{(top) The spurious estimated RVs (black dots) caused by a facula in the simulated data using a Normal and a SN model, the estimated RVs using Eq.~\ref{eq:RV:correction} (red crosses), and the estimated RVs using the usual correction for stellar activity (green triangles), based on $RV_{\text{activity}}=\beta_0+\beta_1 \gamma + \beta_2 \text{SN FWHM}$ for the SN fit and on $RV_{\text{activity}}=\beta_0+\beta_1 \text{BIS SPAN} + \beta_2 \text{FWHM}$ for the Normal fit.    
   %
 (bottom) The residuals from the model fit using Eq.~\ref{eq:RV:correction} (red crosses) and the residuals from the usual correction (green triangles). The std. are also reported in the legend, and the residuals have a smaller systematic component when using the proposed model of Eq.~\ref{eq:RV:correction} compared to the usual model.
The tests of statistical significance on the parameters are presented in Table \ref{table:faculae.test}.
}\label{fig:faculae.correction}
\end{center}
\end{figure*}

\begin{table}
\begin{center}
\caption{P-values for the estimated coefficients from the model in Eq.~\ref{eq:RV:correction} for correcting stellar activity induced by an equatorial 3\% facula on the simulated Sun. All the parameters corresponding to the Normal or SN variables are statistically significant. The estimated $R^{2}$ show that the proposed correction for stellar activity explains the vast majority of the spurious variability present in the different RV estimates.}
\label{table:faculae.test}
\begin{tabular}{|c|c|c|c|}
\hline
Parameter          & N mean RV         &   SN mean RV &   SN median RV \\
\hline
$\beta_{0}$            &    $0.033$    & $0.00020$ & $0.61$ \\
\hline
$\beta_{1}$            &    $2.22e-16 $    & $2.22e-16 $ & $2.22e-16 $ \\
\hline
$\beta_{2}$            &     $0.0034$   &  $2.22e-16 $ & $2.22e-16 $\\
\hline
$\beta_{3}$            &     $0.00016$   &  $1.091e-6$ & $9.75e-7$\\
\hline
$\beta_{4}$            &     $2.22e-16$   &  $2.22e-16$ & $2.22e-16$\\
\hline
$R^{2}$      &     $0.9978$    &  $0.9985$ & $0.9981$  \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

As before, the original RV estimates are corrected using Eq.~\ref{eq:RV:correction}. 
The results of this correction are displayed in Fig.~\ref{fig:spot.correction} and the statistical tests on the coefficients involved in Eq.~\ref{eq:RV:correction} are summarized in Table~\ref{table:spot.test}. 
As with the facula, the proposed correction is able, in the case of the spot, to capture almost entirely the stellar activity signal when considering the SN or Normal parameters, with $R^2$ values for the linear combination larger than 0.99.

In Fig.~\ref{fig:spot.correction}, we see that the proposed activity correction is able to reduce the signal of a spot from a raw RV rms larger than 4.80\ms down to a rms smaller than 0.38 \ms, for any of the different RV estimates.
When comparing the activity correction proposed in this paper with what is commonly used, i.e only a linear combination of the width and asymmetry of the CCF, we see that our solution is capable of reducing the RV residual rms by a factor of 3.5, which is even more than the factor 2 found in the case of the facula.
The Normal or SN parameters involved in Eq.~\ref{eq:RV:correction} are all statistically significant to explain the activity signal as seen in Table~\ref{table:spot.test}, except the width of the CCF when the RVs are derived with the mean or median of the SN density.
%This is not surprising when looking at the circle shape drawn when plotting the width as a function of the RV in Fig.~\ref{fig:spot.corr}.

\begin{figure*}[htbp]
\begin{center}
\includegraphics[height = 6in]{Spot_NEW_CORRECTION_[3]CorrectionActivity_RadialVelocity_vs_time.pdf} 
   \caption{(top) The spurious estimated RVs (black dots) caused by a spot in the simulated data, the estimated RVs using Eq.~\ref{eq:RV:correction} (red crosses) and the estimated RVs using the usual correction for stellar activity (green triangles), based on $RV_{\text{activity}}=\beta_0+\beta_1 \gamma + \beta_2 \text{SN FWHM}$ for the SN fit and on $RV_{\text{activity}}=\beta_0+\beta_1 \text{BIS SPAN} + \beta_2 \text{FWHM}$ for the normal fit.
   %
 (bottom) The residuals from the model fit using Eq.~\ref{eq:RV:correction} (red crosses) and the residuals from the usual correction (green triangles). 
 The std. are also reported in the legend, and the residuals have a smaller systematic component when using the proposed model compared to the usual model.
The tests of statistical significance on the parameters are presented in Table~\ref{table:spot.test}.
}
    \label{fig:spot.correction}
\end{center}
\end{figure*}

\begin{table}
\begin{center}
\caption{P--values for the different coefficients used in Eq.~\ref{eq:RV:correction} for the correction of stellar activity induced by an equatorial 1\% spot on the simulated Sun. All the parameters corresponding to the Normal or SN parameters are statistically significant to explain the spurious RV variations caused by this spot, except for the width of the CCF when using SN mean RV or SN median RV as RV estimates. The estimated $R^{2}$ show that the proposed correction for stellar activity explains the vast majority of the spurious variability seen in the different RV estimates.}
\label{table:spot.test}
\begin{tabular}{|c|c|c|c|}
\hline
Parameter          & N mean RV         &   SN mean RV &   SN median RV \\
\hline
$\beta_{0}$            &    $0.4975$    & $0.21$ & $0.21$ \\
\hline
$\beta_{1}$            &    $2e-16$    & $2e-16$ & $2e-16$ \\
\hline
$\beta_{2}$            &     $2e-16$   &  $2e-16$ & $2e-16$\\
\hline
$\beta_{3}$            &     $0.017$   &  $0.13$ & $0.11$\\
\hline
$\beta_{4}$            &     $2e-16$   &  $2e-16$ & $2e-16$\\
\hline
$R^{2}$      &     $0.9959$    &  $0.9936$ & $0.9952$  \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

In order to correct the estimated RVs from the spurious variation caused by the spot, the proposed model for correcting the activity is added to a planetary signal model that takes into account the RV variations caused by a planet. The observed RVs can therefore be modelled as a combination of the activity and the planetary signals:
%
\begin{equation}
RV= RV_{\text{activity}} + RV_{\text{planet}},
\label{eq:RV:correction.overall}
\end{equation}
%
where $RV_{\text{activity}}$ can be found in Eq.~\ref{eq:RV:correction}, and $RV_{\text{planet}}$, in the case with no eccentricity, can be modelled by the following sinusoidal function:
%
\begin{equation}
RV_{\text{exoplanet}}= K \sin \left(\frac{2 \pi}{P} (t - t_{0})\right),
\label{eq:RV:correction.planet}
\end{equation}
%
with amplitude $K$, orbital period $P$, and an epoch at the periapsis $t_{0}$.  The previous three unknown parameters define the planetary orbit.

%We note that the p--value associated with the amplitude parameter $K$ is particularly relevant for rejecting or not rejecting the assumption about the presence of an orbiting companion. Moreover we note also that Eq.~\ref{eq:RV:correction.planet} is highly non linear, meaning that the estimation of all the parameters involved in Eq.~\ref{eq:RV:correction.overall} has to be done numerically. %We used non linear least squares and the results are displayed in Fig.~\ref{fig:spotplanet.correction}. We can see how in this case we are able to disentangle the spurious variations in RVs caused by stellar activity from the pure Doppler-shift due to the planet. 
The proposed model from Eq.~\ref{eq:RV:correction.overall} was fitted to the RV data and the results of the estimated model are summarized in Table~\ref{table:spotplanet.test}.   Except for the width parameters with coefficient $\beta_3$, all the other Normal or SN parameters are significantly useful to explain the RV variation induced by a spot plus a planet. We also observe that the RV residuals, once corrected for stellar activity and the presence of the planet, are comparable in terms of rms for all the three different RV estimates, with SN mean RV giving a slightly smaller value.
%\begin{figure*}[htbp]
%   \centering
%\includegraphics[height = 4in]{SPOT_PLANET_NEW_CORRECTION.pdf} 
%   \caption{Set of  variations in RVs estimated using a Normal and a SN fit before and once corrected from stellar activity. In this case there are spurious variations caused by the spot and pure Doppler-shift due to the planet. The correction is done using Eq.~\ref{eq:RV:correction.overall} and the estimated parameters are presented in Table \ref{table:spotplanet.test}. In this case, by solving Eq.~\ref{eq:RV:correction.overall}, we are able to completely disentangle the spurious variations in RVs caused by the presence of the spot from the pure dopplershift caused by the exoplanet.}
%    \label{fig:spotplanet.correction}
%\end{figure*}

\begin{table}
\begin{center}
\caption{P-values for the different coefficients used in Eq.~\ref{eq:RV:correction} for the correction of stellar activity induced by an equatorial 1\% spot on the simulated Sun, and a planet with a period of one-third the rotational period of the star and a semi-amplitude  of 10\,\ms. All the parameters corresponding to the Normal or SN variables are statistically significant, except the width of the CCF. Note that since nonlinear least squares was required, the residual standard error rather than the $R^2$ is displayed as a reference.}
\label{table:spotplanet.test}
\begin{tabular}{|c|c|c|c|}
\hline
Parameter          & N mean RV         &   SN mean RV &   SN median RV \\
\hline
$\beta_{0}$            &    $0.00063$    & $2e-16$  & $1.42e-09$ \\
\hline
$\beta_{1}$            &    $2e-16$    & $2e-16$  & $2e-16$ \\
\hline
$\beta_{2}$            &     $2e-16$   & $2e-16$ & $2e-16$\\
\hline
$\beta_{3}$            &     $0.067$   &  $0.40$  & $0.38$\\
\hline
$\beta_{4}$            &     $2e-16$   &  $2e-16$ & $2e-16$\\
\hline
K            &     $2e-16$   &  $2e-16$   & $2e-16$ \\
\hline
P            &     $2e-16$   &  $2e-16$ & $2e-16$ \\
\hline
$t_{0}$            &     $2e-16$   &  $2e-16$ & $2e-16$ \\
\hline
$\text{Residuals}$      &     $0.71 \ms$    &  $ 0.66 \ms$ & $0.70 \ms$  \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

In the previous section, we saw that the estimated RV measured when considering a SN or a Normal density resulted in different amplitudes, especially when using the SN mean RV. 
However, once corrected for stellar activity using the linear combination presented in Eq.~\ref{eq:RV:correction}, as shown in the bottom plots of Fig~\ref{fig:alphacent:correctionRV}, the rms of the residuals are essentially the same for all three approaches.
In this section, we investigate the ability of the three different RV estimators (N mean RV, SN mean RV, and SN median RV) to detect planetary signals among stellar activity, and also compare them when using the usual stellar activity correction with the proposed stellar activity model of Eq.~\eqref{eq:RV:correction}.
To carryout this test, the minimum detected amplitude of an injected planetary signal is estimated at different orbital periods when considering data affected by stellar activity.

>>>>

The Skew Normal (SN) distribution is a class of probability distributions which includes the Normal distribution as a special case \citep{Azzalini1985}. The SN distribution has, in addition to a location and a scale parameter analogous to the Normal distribution's mean and standard deviation, a third parameter which describes the skewness (i.e. the asymmetry) of the distribution. Considering a random variable $Y\in \mathbb R$ (where $\mathbb R$ is the real line) which follows a SN distribution with location parameter $\xi \in \mathbb R$, scale parameter $\omega \in \mathbb R^{+}$ (i.e., the positive real line), and skewness parameter $\alpha \in \mathbb R$, its density at some value $y\in Y$ can be written as 
\begin{equation} \label{def:snd_gen}
SN(y;\xi, \omega, \alpha) = \frac{2}{\omega} \phi\left(\frac{y-\xi}{\omega}\right) \Phi\left(\frac{\alpha(y-\xi)}{\omega}\right),
\end{equation}
where $\phi$ and $\Phi$ are, respectively, the density function and the distribution function of a standard Normal distribution\footnote{A standard Normal distribution is a Normal distribution with a mean of 0 and a standard deviation of 1.}.
The skewness parameter $\alpha$ quantifies the asymmetry of the SN. 
Examples of SN densities under different skewness parameter values and the same location and scale parameters ($\xi = 0$ and $\omega = 1$) are displayed in Fig.~\ref{fig:SN.plot}.  A usual Normal distribution is the special case of the SN distribution when the skewness parameter $\alpha$ is equal to zero\footnote{This can be seen from Eq.~\eqref{def:snd_gen}. If $\alpha = 0$ then $\Phi\left(\frac{\alpha(y-\xi)}{\omega}\right) = \Phi(0) = 0.5$ and therefore $SN(y;\xi, \omega, 0) = \frac{1}{\omega} \phi\left(\frac{y-\xi}{\omega}\right)$ which is the density of a Normal distribution. Note that $\Phi(0) = 0.5$ because $\Phi(0)$ is the the probability that a standard Normal random variable is less than or equal than 0.}.
%
\begin{figure}[t]
\begin{center}
\includegraphics[height = 2.3in]{Skew_Normal_densities_jjck.pdf} 
   \caption{Density function of a random variable $Y$ following the SN distribution $SN(\xi, \omega, \alpha)$ with location parameter $\xi = 0$, scale parameter $\omega = 1$ and different values of the skewness parameter $\alpha$ indicated by different colors and line types. Note that the solid black line has an $\alpha = 0$ making it a Normal distribution.}
   \label{fig:SN.plot}
\end{center}
\end{figure}
%
For reasons related to the interpretation of the parameters in Eq.~\eqref{def:snd_gen} and computational issues with estimating $\alpha$ near 0, a different parametrization is used in this work, which is referred to as the \emph{centered parametrization} (CP).  This CP is much closer to the parametrization of a Normal distribution, as it uses a mean parameter $\mu$, a variance parameter $\sigma^2$ and a skewness parameter $\gamma$. In order to define the CP, we need to express the CP parameters $(\mu, \sigma^2, \gamma)$ as a function of $(\xi, \omega^2, \alpha)$. This can be done using the following relations:
%
\begin{equation} \label{eq:snd_cp}
\mu = \xi + \omega \beta, \quad \sigma^{2} = \omega^{2}(1-\beta^2), \quad \gamma = \frac{1}{2}(4-\pi) \beta^{3}\left(1-\beta^2\right)^{-3/2},
\end{equation}
%
where $\beta = \sqrt{\frac{2}{\pi}}\left(\frac{\alpha}{\sqrt{1+\alpha^2}}\right)$ \citep[e.g.][]{Arellano-2010}.

By using Eq.~\eqref{eq:snd_cp}, the new set of parameters $(\mu, \sigma^2, \gamma)$ provides a clearer interpretation of the behavior of the SN distribution. For the $\alpha$ values used in Fig.~\ref{fig:SN.plot}, the corresponding values of ($\mu$, $\sigma^2$, $\gamma$) are displayed in Table~\ref{tab:cp_values}.  In particular, $\mu$ and $\sigma^2$ are the actual mean and variance of the distribution, rather than simply a location and scale parameter, and $\gamma$ provides a measure of the skewness of the SN. 
Along with the mean of the SN, we consider the median of the distribution as a measure of its barycenter.  See Table~\ref{tab:cp_values} for the medians of the SN densities displayed in Fig.~\ref{fig:SN.plot}.

@@

\end{center}
\end{table}
%
Further details about the parametrization from Eq.~\eqref{def:snd_gen}, called the \emph{Direct Parametrization} (DP), the CP, and general statistical properties of the SN can be found in \cite{Azzalini2014}.

@@

The proposed function defined in Eq.~\eqref{eq:RV:correction} is the result of statistical and astronomical considerations. 
In particular, we considered issues with correlations between the model covariates, which can lead to issues because the matrix needed to calculate the estimates could be singular; this problem is known in statistics as \emph{multicollinearity} and more discussion can be found in \citet{belsley1991}.
In the analysis of real data presented in this work, we never observed a correlation coefficient exceeding $0.66$ between the asymmetry and width parameters and therefore the problem of multicollinearity appeared to be mitigated. 
We also investigated the statistical significant of the interactions between $A$ and the width of the CCF, and $A$ and the asymmetry of the CCF; those two interactions were not statistically significant.

%-----------------------------------------------------------------------------------------------------------------------------------------------
\section{Simulation Study} \label{sec:soap}
In order to evaluate the performance of the proposed SN approach for modelling the CCF and the benefit of using the proposed correction for stellar activity (see Eq.~\eqref{eq:RV:correction}), we begin by considering a simulation study using spectra generated from the Spot Oscillation And Planet 2.0 code \citep[SOAP 2.0,][]{Dumusque-2014b}.

@@

The RV variations displayed in Fig.~\ref{fig:faculae} are caused only by stellar activity, in this case a facula, so we applied the activity correction proposed in Eq.~\eqref{eq:RV:correction} to evaluate the performance of the model in this setting.
The results of this correction are displayed in Fig.~\ref{fig:faculae.correction} and the statistical tests on the coefficients involved in Eq.~\eqref{eq:RV:correction} are summarized in Table \ref{table:faculae.test}. 
The proposed correction for stellar activity is able to account for the majority of the activity signal created by a facula, with a $R^2$ larger than $0.99$. 
In addition, the rms of the different estimates of the RV reduces from about 3 \ms before correcting for stellar activity to values below 0.15 \ms after correcting for stellar activity.
We see a slightly smaller rms after correction when using the SN parameters compared to the Normal parameters.
When comparing the correction proposed in Eq.~\eqref{eq:RV:correction} with what is generally used (i.e. a linear combination of only the asymmetry and width parameters), we see that the proposed correction is able to reduce the rms of the RV residuals by a factor of 2. Looking at the significance of the coefficients in Table~\ref{table:faculae.test}, we observe that all the SN or Normal parameters are relevant for the correction.

\begin{figure*}[htbp]
\begin{center}
\includegraphics[height = 6in]{FACULAE_NEW_CORRECTION_[3]CorrectionActivity_RadialVelocity_vs_time.pdf} 
   \caption{(top) The spurious estimated RVs (black dots) caused by a facula in the simulated data using a Normal and a SN model, the estimated RVs using Eq.~\eqref{eq:RV:correction} (red crosses), and the estimated RVs using the usual correction for stellar activity (green triangles), based on $RV_{\text{activity}}=\beta_0+\beta_1 \gamma + \beta_2 \text{SN FWHM}$ for the SN fit and on $RV_{\text{activity}}=\beta_0+\beta_1 \text{BIS SPAN} + \beta_2 \text{FWHM}$ for the Normal fit.    
   %
 (bottom) The residuals from the model fit using Eq.~\eqref{eq:RV:correction} (red crosses) and the residuals from the usual correction (green triangles). The std. are also reported in the legend, and the residuals have a smaller systematic component when using the proposed model of Eq.~\eqref{eq:RV:correction} compared to the usual model.
The tests of statistical significance on the parameters are presented in Table \ref{table:faculae.test}.
}\label{fig:faculae.correction}
\end{center}
\end{figure*}

\begin{table}
\begin{center}
\caption{P-values for the estimated coefficients from the model in Eq.~\eqref{eq:RV:correction} for correcting stellar activity induced by an equatorial 3\% facula on the simulated Sun. All the parameters corresponding to the Normal or SN variables are statistically significant. The estimated $R^{2}$ show that the proposed correction for stellar activity explains the vast majority of the spurious variability present in the different RV estimates.}
\label{table:faculae.test}
\begin{tabular}{|c|c|c|c|}
\hline
Parameter          & N mean RV         &   SN mean RV &   SN median RV \\
\hline
$\beta_{0}$            &    $0.033$    & $0.00020$ & $0.61$ \\
\hline
$\beta_{1}$            &    $2.22e-16 $    & $2.22e-16 $ & $2.22e-16 $ \\
\hline
$\beta_{2}$            &     $0.0034$   &  $2.22e-16 $ & $2.22e-16 $\\
\hline
$\beta_{3}$            &     $0.00016$   &  $1.091e-6$ & $9.75e-7$\\
\hline
$\beta_{4}$            &     $2.22e-16$   &  $2.22e-16$ & $2.22e-16$\\
\hline
$R^{2}$      &     $0.9978$    &  $0.9985$ & $0.9981$  \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

As before, the original RV estimates are corrected using Eq.~\eqref{eq:RV:correction}. 
The results of this correction are displayed in Fig.~\ref{fig:spot.correction} and the statistical tests on the coefficients involved in Eq.~\eqref{eq:RV:correction} are summarized in Table~\ref{table:spot.test}. 
As with the facula, the proposed correction is able, in the case of the spot, to capture almost entirely the stellar activity signal when considering the SN or Normal parameters, with $R^2$ values for the linear combination larger than 0.99.

In Fig.~\ref{fig:spot.correction}, we see that the proposed activity correction is able to reduce the signal of a spot from a raw RV rms larger than 4.80\ms down to a rms smaller than 0.38 \ms, for any of the different RV estimates.
When comparing the activity correction proposed in this paper with what is commonly used, i.e only a linear combination of the width and asymmetry of the CCF, we see that our solution is capable of reducing the RV residual rms by a factor of 3.5, which is even more than the factor 2 found in the case of the facula.
The Normal or SN parameters involved in Eq.~\eqref{eq:RV:correction} are all statistically significant to explain the activity signal as seen in Table~\ref{table:spot.test}, except the width of the CCF when the RVs are derived with the mean or median of the SN density.
%This is not surprising when looking at the circle shape drawn when plotting the width as a function of the RV in Fig.~\ref{fig:spot.corr}.

\begin{figure*}[htbp]
\begin{center}
\includegraphics[height = 6in]{Spot_NEW_CORRECTION_[3]CorrectionActivity_RadialVelocity_vs_time.pdf} 
   \caption{(top) The spurious estimated RVs (black dots) caused by a spot in the simulated data, the estimated RVs using Eq.~\eqref{eq:RV:correction} (red crosses) and the estimated RVs using the usual correction for stellar activity (green triangles), based on $RV_{\text{activity}}=\beta_0+\beta_1 \gamma + \beta_2 \text{SN FWHM}$ for the SN fit and on $RV_{\text{activity}}=\beta_0+\beta_1 \text{BIS SPAN} + \beta_2 \text{FWHM}$ for the normal fit.
   %
 (bottom) The residuals from the model fit using Eq.~\eqref{eq:RV:correction} (red crosses) and the residuals from the usual correction (green triangles). 
 The std. are also reported in the legend, and the residuals have a smaller systematic component when using the proposed model compared to the usual model.
The tests of statistical significance on the parameters are presented in Table~\ref{table:spot.test}.
}
    \label{fig:spot.correction}
\end{center}
\end{figure*}

\begin{table}
\begin{center}
\caption{P--values for the different coefficients used in Eq.~\eqref{eq:RV:correction} for the correction of stellar activity induced by an equatorial 1\% spot on the simulated Sun. All the parameters corresponding to the Normal or SN parameters are statistically significant to explain the spurious RV variations caused by this spot, except for the width of the CCF when using SN mean RV or SN median RV as RV estimates. The estimated $R^{2}$ show that the proposed correction for stellar activity explains the vast majority of the spurious variability seen in the different RV estimates.}
\label{table:spot.test}
\begin{tabular}{|c|c|c|c|}
\hline
Parameter          & N mean RV         &   SN mean RV &   SN median RV \\
\hline
$\beta_{0}$            &    $0.4975$    & $0.21$ & $0.21$ \\
\hline
$\beta_{1}$            &    $2e-16$    & $2e-16$ & $2e-16$ \\
\hline
$\beta_{2}$            &     $2e-16$   &  $2e-16$ & $2e-16$\\
\hline
$\beta_{3}$            &     $0.017$   &  $0.13$ & $0.11$\\
\hline
$\beta_{4}$            &     $2e-16$   &  $2e-16$ & $2e-16$\\
\hline
$R^{2}$      &     $0.9959$    &  $0.9936$ & $0.9952$  \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

In order to correct the estimated RVs from the spurious variation caused by the spot, the proposed model for correcting the activity is added to a planetary signal model that takes into account the RV variations caused by a planet. The observed RVs can therefore be modelled as a combination of the activity and the planetary signals:
%
\begin{equation}
RV= RV_{\text{activity}} + RV_{\text{planet}},
\label{eq:RV:correction.overall}
\end{equation}
%
where $RV_{\text{activity}}$ can be found in Eq.~\eqref{eq:RV:correction}, and $RV_{\text{planet}}$, in the case with no eccentricity, can be modelled by the following sinusoidal function:
%
\begin{equation}
RV_{\text{exoplanet}}= K \sin \left(\frac{2 \pi}{P} (t - t_{0})\right),
\label{eq:RV:correction.planet}
\end{equation}
%
with amplitude $K$, orbital period $P$, and an epoch at the periapsis $t_{0}$.  The previous three unknown parameters define the planetary orbit.

%We note that the p--value associated with the amplitude parameter $K$ is particularly relevant for rejecting or not rejecting the assumption about the presence of an orbiting companion. Moreover we note also that Eq.~\eqref{eq:RV:correction.planet} is highly non linear, meaning that the estimation of all the parameters involved in Eq.~\eqref{eq:RV:correction.overall} has to be done numerically. %We used non linear least squares and the results are displayed in Fig.~\ref{fig:spotplanet.correction}. We can see how in this case we are able to disentangle the spurious variations in RVs caused by stellar activity from the pure Doppler-shift due to the planet. 
The proposed model from Eq.~\eqref{eq:RV:correction.overall} was fitted to the RV data and the results of the estimated model are summarized in Table~\ref{table:spotplanet.test}.   Except for the width parameters with coefficient $\beta_3$, all the other Normal or SN parameters are significantly useful to explain the RV variation induced by a spot plus a planet. We also observe that the RV residuals, once corrected for stellar activity and the presence of the planet, are comparable in terms of rms for all the three different RV estimates, with SN mean RV giving a slightly smaller value.
%\begin{figure*}[htbp]
%   \centering
%\includegraphics[height = 4in]{SPOT_PLANET_NEW_CORRECTION.pdf} 
%   \caption{Set of  variations in RVs estimated using a Normal and a SN fit before and once corrected from stellar activity. In this case there are spurious variations caused by the spot and pure Doppler-shift due to the planet. The correction is done using Eq.~\eqref{eq:RV:correction.overall} and the estimated parameters are presented in Table \ref{table:spotplanet.test}. In this case, by solving Eq.~\eqref{eq:RV:correction.overall}, we are able to completely disentangle the spurious variations in RVs caused by the presence of the spot from the pure dopplershift caused by the exoplanet.}
%    \label{fig:spotplanet.correction}
%\end{figure*}

\begin{table}
\begin{center}
\caption{P-values for the different coefficients used in Eq.~\eqref{eq:RV:correction} for the correction of stellar activity induced by an equatorial 1\% spot on the simulated Sun, and a planet with a period of one-third the rotational period of the star and a semi-amplitude  of 10\,\ms. All the parameters corresponding to the Normal or SN variables are statistically significant, except the width of the CCF. Note that since nonlinear least squares was required, the residual standard error rather than the $R^2$ is displayed as a reference.}
\label{table:spotplanet.test}
\begin{tabular}{|c|c|c|c|}
\hline
Parameter          & N mean RV         &   SN mean RV &   SN median RV \\
\hline
$\beta_{0}$            &    $0.00063$    & $2e-16$  & $1.42e-09$ \\
\hline
$\beta_{1}$            &    $2e-16$    & $2e-16$  & $2e-16$ \\
\hline
$\beta_{2}$            &     $2e-16$   & $2e-16$ & $2e-16$\\
\hline
$\beta_{3}$            &     $0.067$   &  $0.40$  & $0.38$\\
\hline
$\beta_{4}$            &     $2e-16$   &  $2e-16$ & $2e-16$\\
\hline
K            &     $2e-16$   &  $2e-16$   & $2e-16$ \\
\hline
P            &     $2e-16$   &  $2e-16$ & $2e-16$ \\
\hline
$t_{0}$            &     $2e-16$   &  $2e-16$ & $2e-16$ \\
\hline
$\text{Residuals}$      &     $0.71 \ms$    &  $ 0.66 \ms$ & $0.70 \ms$  \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

In the previous section, we saw that the estimated RV measured when considering a SN or a Normal density resulted in different amplitudes, especially when using the SN mean RV. 
However, once corrected for stellar activity using the linear combination presented in Eq.~\eqref{eq:RV:correction}, as shown in the bottom plots of Fig~\ref{fig:alphacent:correctionRV}, the rms of the residuals are essentially the same for all three approaches.
In this section, we investigate the ability of the three different RV estimators (N mean RV, SN mean RV, and SN median RV) to detect planetary signals among stellar activity, and also compare them when using the usual stellar activity correction with the proposed stellar activity model of Eq.~\eqref{eq:RV:correction}.
To carryout this test, the minimum detected amplitude of an injected planetary signal is estimated at different orbital periods when considering data affected by stellar activity.
